---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

## sparseNumeric

<!-- badges: start -->
<!-- badges: end -->

The goal of 'sparseNumeric' is to provide an S4 class for representing sparse numeric vectors - that is, vectors that contain many zeros - and to turn the sparse-vector code from HW5 into a complete, well-documented R package. 

This package stors sparse vectors efficiently by keeping only the non-zero values and their positions. It supports arithmetic operations (ex. addition, multiplication, etc.) and dot product directly on the sparse representation, so the full vector never needs to be rebulit. 

In addition to basic arithmetic, the package also includes useful statistical methods:

- 'mean()' - computes the mean of the vector, counding all zeros, without converting the vector back to dense form.
- 'norm()' - computes the Euclidean norm (the square root of the sum of squared values)
- 'standardize()' - standardizes the vector by subtracting its mean and dividing by its standard deviation. 

All methods operate **directly on the sparse structure**, making them both efficient and memory-friendly. The package includes roxygen2 documentation unit tests, a proper DESCRIPTION file, and Github Actions to ensure the package passes R CMD check. 

Along with these methods, the package includes:

- roxygen2 documentation for the class and all methods  
- a clean DESCRIPTION file with real metadata  
- unit tests written with **testthat**, reaching at least **90% coverage**  
- a passing **R CMD check**  
- a GitHub Actions workflow to run checks on every push  
- a `pkgdown` website documenting all functions  

## Installation

You can install the development version of sparseNumeric from Github:

``` r
# install.packages("devtools")
devtools::install_github("algodines/HW6_sparce_numeric")
```

## Example

This is a basic example which shows you how to solve a common problem:

```{r example}
library(sparseNumeric)

# Create a regular R vector
v <- c(0, 0, 5, 0, 3)

# Convert to sparse_numeric
s <- as(v, "sparse_numeric")

# Display the sparse object
s

# Compute statistics
mean(s)
norm(s)
standardize(s)
```

You can also perform arithemtic using the sparse methods:

```{r}
x <- as(c(0, 1, 0, 2), "sparse_numeric")
y <- as(c(3, 0, 4, 0), "sparse_numeric")

x + y     # sparse addition
x - y     # sparse subtraction
x * y     # elementwise product
sparse_crossprod(x, y)  # dot product
```

Why use README.Rmd?
'`README.Rmd` is helpful because:
- you can include R code chunks that run automatically
- output (numbers, tables,plots) appears directly in your README
- your README stays in sync with your actual code
- you can regenerate it at anytime with:

```r
devtools::build_readme()
```

For example, here is a small built-in plot:

```{r cars}
summary(cars)
```

You can also embed plots, for example:

```{r pressure, echo = FALSE}
plot(pressure)
```
